Supplementary Material for `The response of trophic interaction networks to multiple stressors in a marine latitudinal gradient of the Southern Hemisphere'

### 2024-03-25 manuscript.tex
Change to Rmd for trackdow

Here we review for the first time the state-of-the-art knowledge on
stressor response of marine food webs along the southwest Atlantic to
Antarctic gradient (45 - 78ºS, Figure 1). We focus on proven and
expected changes in food webs driven by stressors in selected areas
along this large-scale latitudinal gradient. We recruited food web
experts from different marine systems of Argentina and the world (see
co-authors' list). Throughout the year 2023 we maintained regular
discussion meetings, typically held every two or three weeks. The aim of
this review is threefold: 1) describe the complexity and structure of
marine food webs along the southwest Atlantic to Antarctic axis from a
network perspective; 2) identify the ongoing environmental and
anthropogenic stressors for each marine ecosystem containing the food
webs; and 3) resume proven food web changes and elaborate hypotheses on
how the identified stressors might affect food web features (e.g.~energy
flow, stability), combining information on node- and network-level
properties. Finally, we suggest which additional data and analyses are
needed to gain insights into the stressors' effects on food web
properties in the southwest Atlantic to Antarctic region.

@@

\bibitem[\citeproctext]{ref-APN2022}
Administración de Parques Nacionales. 2022. {Plan de gestión AMP
Namuncurá Banco Burdwood}. {Dirección Nacional de Áreas Marinas
Protegidas (DNAMP), Argentina}.

\bibitem[\citeproctext]{ref-Albert2002}
Albert, R., and Barabási, A.-L. 2002. Statistical mechanics of complex
networks. Reviews of Modern Physics \textbf{74}(1): 47--97. {American
Physical Society}.
doi:\href{https://doi.org/10.1103/RevModPhys.74.47}{10.1103/RevModPhys.74.47}.

\bibitem[\citeproctext]{ref-Allega2020}
Allega, L., Braverman, M., Campodónico, S., Carozza, C.R., Cepeda, G.D.,
Colonello, J.H., Derisio, C., Di Mauro, R., Firpo, C.A., Gaitán, E.,
Hozbor, C., Irusta, G., Ivanovic, M., Lagos, N., Lutz, V.A., Marí, N.R.,
Militelli, M.I., Moriondo Danovaro, P., Navarro, G., Orlando, P.,
Pájaro, M., Prandoni, N., Prosdocimi, L., Reta, R., Rico, M.R., Riestra,
C., Ruarte, C., Schejter, L., Schiariti, A., Segura, V., Souto, V.S.,
Temperoni, B., and Verón, E. 2020. {Estado del conocimiento biológico
pesquero de los principales recursos vivos y su ambiente, con relación a
la exploración hidrocarburífera en la Zona Económica Exclusiva argentina
y sus adyacencias}. {Instituto Nacional de Investigación y Desarrollo
Pesquero (INIDEP)}, {Mar del Plata}. {[}accessed 26 September 2023{]}.

\bibitem[\citeproctext]{ref-Allega2019}
Allega, L., Braverman, M.S., Cabreira, A.G., Campodónico, S., Colonello,
J.H., Derisio, C.M., Di Mauro, R.P., Gaitán, E.N., Hozbor, M., Irusta,
G., Lutz, V.A., Marí, N., Militelli, M.I., Moriondo, P., Navarro, G.,
Orlando, P., Pajaro, M., Prandoni García, N.I., Prosdocimi, L., Reta,
R., Rico, R., Riestra, C.M., Segura, V., Schejter, L., Schiariti, A.,
Souto, V., and Verón, E.M. 2019. {Estado del conocimiento biológico
pesquero de los principales recursos vivos y su ambiente, con relación a
la exploración hidrocarburífera en la Zona Económica Exclusiva Argentina
y adyacencias}. {INIDEP}. {[}accessed 10 August 2023{]}.

\bibitem[\citeproctext]{ref-Anton2019}
Anton, A., Geraldi, N.R., Lovelock, C.E., Apostolaki, E.T., Bennett, S.,
Cebrian, J., Krause-Jensen, D., Marbà, N., Martinetto, P., Pandolfi,
J.M., Santana-Garcon, J., and Duarte, C.M. 2019. Global ecological
impacts of marine exotic species. Nature Ecology \& Evolution
\textbf{3}(5): 787--800. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41559-019-0851-0}{10.1038/s41559-019-0851-0}.

@@

\bibitem[\citeproctext]{ref-Atkinson2019}
Atkinson, A., Hill, S.L., Pakhomov, E.A., Siegel, V., Reiss, C.S., Loeb,
V.J., Steinberg, D.K., Schmidt, K., Tarling, G.A., Gerrish, L., and
Sailley, S.F. 2019. Krill ({Euphausia} superba) distribution contracts
southward during rapid regional warming. Nature Climate Change
\textbf{9}(2): 142--147. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41558-018-0370-z}{10.1038/s41558-018-0370-z}.

\bibitem[\citeproctext]{ref-Atkinson2004}
Atkinson, A., Siegel, V., Pakhomov, E., and Rothery, P. 2004. Long-term
decline in krill stock and increase in salps within the {Southern
Ocean}. Nature \textbf{432}(7013): 100--103. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/nature02996}{10.1038/nature02996}.

\bibitem[\citeproctext]{ref-Balech2008}
Balech, E., and Ehrlich, M.D. 2008. {Esquema biogeográfico del Mar
Argentino}. {Mar del Plata: Instituto Nacional de Investigación y
Desarrollo Pesquero (INIDEP)}. {[}accessed 14 August 2023{]}.

\bibitem[\citeproctext]{ref-Bartley2019}
Bartley, T.J., McCann, K.S., Bieg, C., Cazelles, K., Granados, M.,
Guzzo, M.M., MacDougall, A.S., Tunney, T.D., and McMeans, B.C. 2019.
Food web rewiring in a changing world. Nature Ecology \& Evolution
\textbf{3}(3): 345--354. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41559-018-0772-3}{10.1038/s41559-018-0772-3}.

@@

\bibitem[\citeproctext]{ref-Belgrano2005}
Belgrano, A., Scharler, S.E.R.C.U.M., Scharler, U.M., Dunne, J., and
Ulanowicz, R.E. 2005. Aquatic {Food Webs}: {An Ecosystem Approach}. {OUP
Oxford}.

@@

\bibitem[\citeproctext]{ref-Bovcon2013}
Bovcon, N.D., Góngora, M.E., Marinao, C., and González-Zevallos, D.
2013. Composición de las capturas y descartes generados en la pesca de
merluza común {Merluccius} hubbsi y langostino patagónico {Pleoticus}
muelleri: Un caso de estudio en la flota fresquera de altura del {Golfo
San Jorge}, {Chubut}, {Argentina}. Revista de biología marina y
oceanografía \textbf{48}(2): 303--319. {Universidad de Valparaíso.
Facultad de Ciencias del Mar}.
doi:\href{https://doi.org/10.4067/S0718-19572013000200010}{10.4067/S0718-19572013000200010}.

\bibitem[\citeproctext]{ref-Boyd2015}
Boyd, P.W., Lennartz, S.T., Glover, D.M., and Doney, S.C. 2015.
Biological ramifications of climate-change-mediated oceanic
multi-stressors. Nature Climate Change \textbf{5}(1): 71--79. {Nature
Publishing Group}.
doi:\href{https://doi.org/10.1038/nclimate2441}{10.1038/nclimate2441}.

@@

\bibitem[\citeproctext]{ref-Cirtwill2015}
Cirtwill, A.R., Stouffer, D.B., and Romanuk, T.N. 2015. Latitudinal
gradients in biotic niche breadth vary across ecosystem types.
Proceedings of the Royal Society B: Biological Sciences
\textbf{282}(1819): 20151589. {Royal Society}.
doi:\href{https://doi.org/10.1098/rspb.2015.1589}{10.1098/rspb.2015.1589}.

\bibitem[\citeproctext]{ref-Collins2010}
Collins, M.A., Brickle, P., Brown, J., and Belchier, M. 2010. Chapter
{Four} - {The Patagonian Toothfish}: {Biology}, {Ecology} and {Fishery}.
\emph{In} Advances in {Marine Biology}. \emph{Edited by} M. Lesser.
{Academic Press}. pp. 227--300.
doi:\href{https://doi.org/10.1016/B978-0-12-381015-1.00004-6}{10.1016/B978-0-12-381015-1.00004-6}.

@@

\bibitem[\citeproctext]{ref-Cossi2021}
Cossi, P.F., Ojeda, M., Chiesa, I.L., Rimondino, G.N., Fraysse, C.,
Calcagno, J., and Pérez, A.F. 2021. First evidence of microplastics in
the {Marine Protected Area Namuncurá} at {Burdwood Bank}, {Argentina}: A
study on {Henricia} obesa and {Odontaster} penicillatus
({Echinodermata}: {Asteroidea}). Polar Biology \textbf{44}(12):
2277--2287.
doi:\href{https://doi.org/10.1007/s00300-021-02959-5}{10.1007/s00300-021-02959-5}.

\bibitem[\citeproctext]{ref-Cote2016}
Côté, I.M., Darling, E.S., and Brown, C.J. 2016. Interactions among
ecosystem stressors and their importance in conservation. Proceedings of
the Royal Society B: Biological Sciences \textbf{283}(1824): 20152592.
{Royal Society}.
doi:\href{https://doi.org/10.1098/rspb.2015.2592}{10.1098/rspb.2015.2592}.

\bibitem[\citeproctext]{ref-deHaro2022}
de Haro, J.C., Perez Orsi, H., Cané, S., Di Pangracio, A., Falabella,
V., and Sapoznikow, A. 2022. {Informe colaborativo sobre el Estado de
situación. Riesgos e impactos de la prospección sísmica en el Mar
Argentino}. {Foro para la Conservación del Mar Patagónico y Áreas de
Influencia}.

@@

\bibitem[\citeproctext]{ref-Deregibus2017}
Deregibus, D., Quartino, M.L., Zacher, K., Campana, G.L., and Barnes,
D.K.A. 2017. Understanding the link between sea ice, ice scour and
{Antarctic} benthic biodiversity\textendash the need for cross-station
and international collaboration. Polar Record \textbf{53}(2): 143--152.
doi:\href{https://doi.org/10.1017/S0032247416000875}{10.1017/S0032247416000875}.

\bibitem[\citeproctext]{ref-Deregibus2020}
Deregibus, D., Zacher, K., Bartsch, I., Campana, G.L., Momo, F.R.,
Wiencke, C., Gómez, I., and Quartino, M.L. 2020. Carbon {Balance Under}
a {Changing Light Environment}. \emph{In} Antarctic {Seaweeds}:
{Diversity}, {Adaptation} and {Ecosystem Services}. \emph{Edited by} I.
Gómez and P. Huovinen. {Springer International Publishing}, {Cham}. pp.
173--191.
doi:\href{https://doi.org/10.1007/978-3-030-39448-6_9}{10.1007/978-3-030-39448-6\_9}.

@@

\bibitem[\citeproctext]{ref-Dulvy2014}
Dulvy, N.K., Fowler, S.L., Musick, J.A., Cavanagh, R.D., Kyne, P.M.,
Harrison, L.R., Carlson, J.K., Davidson, L.N., Fordham, S.V., Francis,
M.P., Pollock, C.M., Simpfendorfer, C.A., Burgess, G.H., Carpenter,
K.E., Compagno, L.J., Ebert, D.A., Gibson, C., Heupel, M.R.,
Livingstone, S.R., Sanciangco, J.C., Stevens, J.D., Valenti, S., and
White, W.T. 2014. Extinction risk and conservation of the world's sharks
and rays. eLife \textbf{3}: e00590. {eLife Sciences Publications, Ltd}.
doi:\href{https://doi.org/10.7554/eLife.00590}{10.7554/eLife.00590}.

@@

\bibitem[\citeproctext]{ref-Dunne2002a}
Dunne, J.A., Williams, R.J., and Martinez, N.D. 2002b. Food-web
structure and network theory: {The} role of connectance and size.
Proceedings of the National Academy of Sciences \textbf{99}(20):
12917--12922. {Proceedings of the National Academy of Sciences}.
doi:\href{https://doi.org/10.1073/pnas.192407699}{10.1073/pnas.192407699}.

@@

\bibitem[\citeproctext]{ref-Fretwell2023}
Fretwell, P.T., Boutet, A., and Ratcliffe, N. 2023. Record low 2022
{Antarctic} sea ice led to catastrophic breeding failure of emperor
penguins. Communications Earth \& Environment \textbf{4}(1): 1--6.
{Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s43247-023-00927-x}{10.1038/s43247-023-00927-x}.

\bibitem[\citeproctext]{ref-Fuentes2016}
Fuentes, V., Alurralde, G., Meyer, B., Aguirre, G.E., Canepa, A., Wölfl,
A.-C., Hass, H.C., Williams, G.N., and Schloss, I.R. 2016. Glacial
melting: An overlooked threat to {Antarctic} krill. Scientific Reports
\textbf{6}(1): 27234. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/srep27234}{10.1038/srep27234}.

\bibitem[\citeproctext]{ref-Funes2020}
Funes, M. 2020, March. Efectos de la pesca de arrastre sobre la
estructura trófica del norte del {Golfo San Jorge}. PhD thesis.

\bibitem[\citeproctext]{ref-Funes2019}
Funes, M., Marinao, C., and Galván, D.E. 2019. Does trawl fisheries
affect the diet of fishes? {A} stable isotope analysis approach*.
Isotopes in Environmental and Health Studies \textbf{55}(4): 327--343.
{Taylor \& Francis}.
doi:\href{https://doi.org/10.1080/10256016.2019.1626381}{10.1080/10256016.2019.1626381}.

\bibitem[\citeproctext]{ref-Funes2022}
Funes, M., Saravia, L.A., Cordone, G., Iribarne, O.O., and Galván, D.E.
2022. Network analysis suggests changes in food web stability produced
by bottom trawl fishery in {Patagonia}. Scientific Reports
\textbf{12}(1): 10876. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41598-022-14363-y}{10.1038/s41598-022-14363-y}.

\bibitem[\citeproctext]{ref-Gaitan2016}
Gaitán, E., and Marí, N. 2016. {Análisis de las comunidades bentónicas
asociadas a capturas de la flota comercial dirigida a Macruronus
magellanicus}. {Instituto Nacional de Investigación y Desarrollo
Pesquero (INIDEP)}.

\bibitem[\citeproctext]{ref-Galvan2022}
Galván, D.E., Bovcon, N.D., Cochia, P.D., González, R.A., Lattuca, M.E.,
Reinaldo, M.O., Rincón-Díaz, M.P., Romero, M.A., Vanella, F.A., Venerus,
L.A., and Svendsen, G.M. 2022. Changes in the {Specific} and
{Biogeographic Composition} of {Coastal Fish Assemblages} in
{Patagonia}, {Driven} by {Climate Change}, {Fishing}, and {Invasion} by
{Alien Species}. \emph{In} Global {Change} in {Atlantic Coastal
Patagonian Ecosystems}: {A Journey Through Time}. \emph{Edited by} E.W.
Helbling, M.A. Narvarte, R.A. González, and V.E. Villafañe. {Springer
International Publishing}, {Cham}. pp. 205--231.
doi:\href{https://doi.org/10.1007/978-3-030-86676-1_9}{10.1007/978-3-030-86676-1\_9}.

@@

\bibitem[\citeproctext]{ref-Garcia-Borboroglu2008}
García-Borboroglu, P., Boersma, P.D., Reyes, L., and Skewgar, E. 2008.
Petroleum {Pollution} and {Penguins}: {Marine Conservation Tools} to
{Reduce} the {Problem}. \emph{In} Marine {Pollution}: {New} research,
Hofer, T.N. {Nova Science Publishers Inc.}, {New York, USA}. pp.
339--356.

@@

\bibitem[\citeproctext]{ref-Gellner2023}
Gellner, G., McCann, K., and Hastings, A. 2023. Stable diverse food webs
become more common when interactions are more biologically constrained.
Proceedings of the National Academy of Sciences \textbf{120}(31):
e2212061120. {Proceedings of the National Academy of Sciences}.
doi:\href{https://doi.org/10.1073/pnas.2212061120}{10.1073/pnas.2212061120}.

@@

\bibitem[\citeproctext]{ref-Gongora2012}
Góngora, M.E., González-Zevallos, D., Pettovello, A., and Mendía, L.
2012. Caracterización de las principales pesquerías del golfo {San Jorge
Patagonia}, {Argentina}. Latin american journal of aquatic research
\textbf{40}(1): 1--11. {Pontificia Universidad Católica de Valparaíso.
Facultad de Recursos Naturales. Escuela de Ciencias del Mar}.
{[}accessed 14 August 2023{]}.

@@

\bibitem[\citeproctext]{ref-Gorini2021}
Gorini, F.L., Lukaszewicz, G., and Giussi, A.R. 2021. {Actualización de
la estadística pesquera de peces demersales australes en el Atlántico
Sudoccidental (período 2008-2020)}. \{Informe
T\textbackslash\textquotesingle ecnico Oficial\}, {Instituto Nacional de
Investigación y Desarrollo Pesquero (INIDEP)}, {Mar del Plata}.

\bibitem[\citeproctext]{ref-Guihou2020}
Guihou, K., Piola, A.R., Palma, E.D., and Chidichimo, M.P. 2020.
Dynamical connections between large marine ecosystems of austral {South
America} based on numerical simulations. Ocean Science \textbf{16}(2):
271--290. {Copernicus GmbH}.
doi:\href{https://doi.org/10.5194/os-16-271-2020}{10.5194/os-16-271-2020}.

@@

\bibitem[\citeproctext]{ref-Kawaguchi2009}
Kawaguchi, S., Nicol, S., and Press, A. 2009. Direct effects of climate
change on the {Antarctic} krill fishery. {University Of Tasmania}.
{[}accessed 13 September 2023{]}.

>>>>

Here we review for the first time the state-of-the-art knowledge on
stressor response of marine food webs along the southwest Atlantic to
Antarctic gradient (45 - 78ºS, Figure 1). We focus on proven and
expected changes in food webs driven by stressors in selected areas
along this large-scale latitudinal gradient. We recruited food web
experts from different marine systems of Argentina and the world (see
co-authors' list). Throughout the year 2023 we maintained regular
discussion meetings, typically held every two or three weeks. The aim of
this review is threefold: 1) describe the complexity and structure of
marine food webs along the southwest Atlantic to Antarctic axis from a
network perspective; 2) identify the ongoing environmental and
anthropogenic stressors for each marine ecosystem containing the food
webs; and 3) review proven food web changes and elaborate hypotheses on
how the identified stressors might affect food web features (e.g.~energy
flow, stability), combining information on node- and network-level
properties. Finally, we suggest which additional data and analyses are
needed to gain insights into the stressors' effects on food web
properties in the southwest Atlantic to Antarctic region.

@@

\bibitem[\citeproctext]{ref-APN2022}
Administración de Parques Nacionales. 2022. {Plan de gesti{ó}n AMP
Namuncur{á} Banco Burdwood}. Direcci{ó}n Nacional de {Á}reas Marinas
Protegidas (DNAMP), Argentina.

\bibitem[\citeproctext]{ref-Albert2002}
Albert, R., and Barabási, A.-L. 2002. Statistical mechanics of complex
networks. Reviews of Modern Physics \textbf{74}(1): 47--97. American
Physical Society.
doi:\href{https://doi.org/10.1103/RevModPhys.74.47}{10.1103/RevModPhys.74.47}.

\bibitem[\citeproctext]{ref-Allega2020}
Allega, L., Braverman, M., Campodónico, S., Carozza, C.R., Cepeda, G.D.,
Colonello, J.H., Derisio, C., Di Mauro, R., Firpo, C.A., Gaitán, E.,
Hozbor, C., Irusta, G., Ivanovic, M., Lagos, N., Lutz, V.A., Marí, N.R.,
Militelli, M.I., Moriondo Danovaro, P., Navarro, G., Orlando, P.,
Pájaro, M., Prandoni, N., Prosdocimi, L., Reta, R., Rico, M.R., Riestra,
C., Ruarte, C., Schejter, L., Schiariti, A., Segura, V., Souto, V.S.,
Temperoni, B., and Verón, E. 2020. {Estado del conocimiento biol{ó}gico
pesquero de los principales recursos vivos y su ambiente, con relaci{ó}n
a la exploraci{ó}n hidrocarbur{í}fera en la Zona Econ{ó}mica Exclusiva
argentina y sus adyacencias}. Instituto Nacional de Investigaci{ó}n y
Desarrollo Pesquero (INIDEP), Mar del Plata. {[}accessed 26 September
2023{]}.

\bibitem[\citeproctext]{ref-Allega2019}
Allega, L., Braverman, M.S., Cabreira, A.G., Campodónico, S., Colonello,
J.H., Derisio, C.M., Di Mauro, R.P., Gaitán, E.N., Hozbor, M., Irusta,
G., Lutz, V.A., Marí, N., Militelli, M.I., Moriondo, P., Navarro, G.,
Orlando, P., Pajaro, M., Prandoni García, N.I., Prosdocimi, L., Reta,
R., Rico, R., Riestra, C.M., Segura, V., Schejter, L., Schiariti, A.,
Souto, V., and Verón, E.M. 2019. {Estado del conocimiento biol{ó}gico
pesquero de los principales recursos vivos y su ambiente, con relaci{ó}n
a la exploraci{ó}n hidrocarbur{í}fera en la Zona Econ{ó}mica Exclusiva
Argentina y adyacencias}. INIDEP. {[}accessed 10 August 2023{]}.

\bibitem[\citeproctext]{ref-Anton2019}
Anton, A., Geraldi, N.R., Lovelock, C.E., Apostolaki, E.T., Bennett, S.,
Cebrian, J., Krause-Jensen, D., Marbà, N., Martinetto, P., Pandolfi,
J.M., Santana-Garcon, J., and Duarte, C.M. 2019. Global ecological
impacts of marine exotic species. Nature Ecology \& Evolution
\textbf{3}(5): 787--800. Nature Publishing Group.
doi:\href{https://doi.org/10.1038/s41559-019-0851-0}{10.1038/s41559-019-0851-0}.

@@

\bibitem[\citeproctext]{ref-Atkinson2019}
Atkinson, A., Hill, S.L., Pakhomov, E.A., Siegel, V., Reiss, C.S., Loeb,
V.J., Steinberg, D.K., Schmidt, K., Tarling, G.A., Gerrish, L., and
Sailley, S.F. 2019. Krill ({Euphausia} superba) distribution contracts
southward during rapid regional warming. Nature Climate Change
\textbf{9}(2): 142--147. Nature Publishing Group.
doi:\href{https://doi.org/10.1038/s41558-018-0370-z}{10.1038/s41558-018-0370-z}.

\bibitem[\citeproctext]{ref-Atkinson2004}
Atkinson, A., Siegel, V., Pakhomov, E., and Rothery, P. 2004. Long-term
decline in krill stock and increase in salps within the {Southern
Ocean}. Nature \textbf{432}(7013): 100--103. Nature Publishing Group.
doi:\href{https://doi.org/10.1038/nature02996}{10.1038/nature02996}.

\bibitem[\citeproctext]{ref-Balech2008}
Balech, E., and Ehrlich, M.D. 2008. {Esquema biogeogr{á}fico del Mar
Argentino}. Mar del Plata: Instituto Nacional de Investigaci{ó}n y
Desarrollo Pesquero (INIDEP). {[}accessed 14 August 2023{]}.

\bibitem[\citeproctext]{ref-Bartley2019}
Bartley, T.J., McCann, K.S., Bieg, C., Cazelles, K., Granados, M.,
Guzzo, M.M., MacDougall, A.S., Tunney, T.D., and McMeans, B.C. 2019.
Food web rewiring in a changing world. Nature Ecology \& Evolution
\textbf{3}(3): 345--354. Nature Publishing Group.
doi:\href{https://doi.org/10.1038/s41559-018-0772-3}{10.1038/s41559-018-0772-3}.

@@

\bibitem[\citeproctext]{ref-Belgrano2005}
Belgrano, A., Scharler, S.E.R.C.U.M., Scharler, U.M., Dunne, J., and
Ulanowicz, R.E. 2005. Aquatic {Food Webs}: {An Ecosystem Approach}. OUP
Oxford.

@@

\bibitem[\citeproctext]{ref-Bovcon2013}
Bovcon, N.D., Góngora, M.E., Marinao, C., and González-Zevallos, D.
2013. Composici{ó}n de las capturas y descartes generados en la pesca de
merluza com{ú}n {Merluccius} hubbsi y langostino patag{ó}nico
{Pleoticus} muelleri: Un caso de estudio en la flota fresquera de altura
del {Golfo San Jorge}, {Chubut}, {Argentina}. Revista de biolog{í}a
marina y oceanograf{í}a \textbf{48}(2): 303--319. Universidad de
Valpara{í}so. Facultad de Ciencias del Mar.
doi:\href{https://doi.org/10.4067/S0718-19572013000200010}{10.4067/S0718-19572013000200010}.

\bibitem[\citeproctext]{ref-Boyd2015}
Boyd, P.W., Lennartz, S.T., Glover, D.M., and Doney, S.C. 2015.
Biological ramifications of climate-change-mediated oceanic
multi-stressors. Nature Climate Change \textbf{5}(1): 71--79. Nature
Publishing Group.
doi:\href{https://doi.org/10.1038/nclimate2441}{10.1038/nclimate2441}.

@@

\bibitem[\citeproctext]{ref-Cirtwill2015}
Cirtwill, A.R., Stouffer, D.B., and Romanuk, T.N. 2015. Latitudinal
gradients in biotic niche breadth vary across ecosystem types.
Proceedings of the Royal Society B: Biological Sciences
\textbf{282}(1819): 20151589. Royal Society.
doi:\href{https://doi.org/10.1098/rspb.2015.1589}{10.1098/rspb.2015.1589}.

\bibitem[\citeproctext]{ref-Collins2010}
Collins, M.A., Brickle, P., Brown, J., and Belchier, M. 2010. Chapter
{Four} - {The Patagonian Toothfish}: {Biology}, {Ecology} and {Fishery}.
\emph{In} Advances in {Marine Biology}. \emph{Edited by} M. Lesser.
Academic Press. pp. 227--300.
doi:\href{https://doi.org/10.1016/B978-0-12-381015-1.00004-6}{10.1016/B978-0-12-381015-1.00004-6}.

@@

\bibitem[\citeproctext]{ref-Cossi2021}
Cossi, P.F., Ojeda, M., Chiesa, I.L., Rimondino, G.N., Fraysse, C.,
Calcagno, J., and Pérez, A.F. 2021. First evidence of microplastics in
the {Marine Protected Area Namuncur{á}} at {Burdwood Bank}, {Argentina}:
A study on {Henricia} obesa and {Odontaster} penicillatus
({Echinodermata}: {Asteroidea}). Polar Biology \textbf{44}(12):
2277--2287.
doi:\href{https://doi.org/10.1007/s00300-021-02959-5}{10.1007/s00300-021-02959-5}.

\bibitem[\citeproctext]{ref-Cote2016}
Côté, I.M., Darling, E.S., and Brown, C.J. 2016. Interactions among
ecosystem stressors and their importance in conservation. Proceedings of
the Royal Society B: Biological Sciences \textbf{283}(1824): 20152592.
Royal Society.
doi:\href{https://doi.org/10.1098/rspb.2015.2592}{10.1098/rspb.2015.2592}.

\bibitem[\citeproctext]{ref-deHaro2022}
de Haro, J.C., Perez Orsi, H., Cané, S., Di Pangracio, A., Falabella,
V., and Sapoznikow, A. 2022. {Informe colaborativo sobre el Estado de
situaci{ó}n. Riesgos e impactos de la prospecci{ó}n s{í}smica en el Mar
Argentino}. Foro para la Conservaci{ó}n del Mar Patag{ó}nico y {Á}reas
de Influencia.

@@

\bibitem[\citeproctext]{ref-Deregibus2017}
Deregibus, D., Quartino, M.L., Zacher, K., Campana, G.L., and Barnes,
D.K.A. 2017. Understanding the link between sea ice, ice scour and
{Antarctic} benthic biodiversity--the need for cross-station and
international collaboration. Polar Record \textbf{53}(2): 143--152.
doi:\href{https://doi.org/10.1017/S0032247416000875}{10.1017/S0032247416000875}.

\bibitem[\citeproctext]{ref-Deregibus2020}
Deregibus, D., Zacher, K., Bartsch, I., Campana, G.L., Momo, F.R.,
Wiencke, C., Gómez, I., and Quartino, M.L. 2020. Carbon {Balance Under}
a {Changing Light Environment}. \emph{In} Antarctic {Seaweeds}:
{Diversity}, {Adaptation} and {Ecosystem Services}. \emph{Edited by} I.
Gómez and P. Huovinen. Springer International Publishing, Cham. pp.
173--191.
doi:\href{https://doi.org/10.1007/978-3-030-39448-6_9}{10.1007/978-3-030-39448-6\_9}.

@@

\bibitem[\citeproctext]{ref-Dulvy2014}
Dulvy, N.K., Fowler, S.L., Musick, J.A., Cavanagh, R.D., Kyne, P.M.,
Harrison, L.R., Carlson, J.K., Davidson, L.N., Fordham, S.V., Francis,
M.P., Pollock, C.M., Simpfendorfer, C.A., Burgess, G.H., Carpenter,
K.E., Compagno, L.J., Ebert, D.A., Gibson, C., Heupel, M.R.,
Livingstone, S.R., Sanciangco, J.C., Stevens, J.D., Valenti, S., and
White, W.T. 2014. Extinction risk and conservation of the world's sharks
and rays. eLife \textbf{3}: e00590. eLife Sciences Publications, Ltd.
doi:\href{https://doi.org/10.7554/eLife.00590}{10.7554/eLife.00590}.

@@

\bibitem[\citeproctext]{ref-Dunne2002a}
Dunne, J.A., Williams, R.J., and Martinez, N.D. 2002b. Food-web
structure and network theory: {The} role of connectance and size.
Proceedings of the National Academy of Sciences \textbf{99}(20):
12917--12922. Proceedings of the National Academy of Sciences.
doi:\href{https://doi.org/10.1073/pnas.192407699}{10.1073/pnas.192407699}.

@@

\bibitem[\citeproctext]{ref-Fretwell2023}
Fretwell, P.T., Boutet, A., and Ratcliffe, N. 2023. Record low 2022
{Antarctic} sea ice led to catastrophic breeding failure of emperor
penguins. Communications Earth \& Environment \textbf{4}(1): 1--6.
Nature Publishing Group.
doi:\href{https://doi.org/10.1038/s43247-023-00927-x}{10.1038/s43247-023-00927-x}.

\bibitem[\citeproctext]{ref-Fuentes2016}
Fuentes, V., Alurralde, G., Meyer, B., Aguirre, G.E., Canepa, A., Wölfl,
A.-C., Hass, H.C., Williams, G.N., and Schloss, I.R. 2016. Glacial
melting: An overlooked threat to {Antarctic} krill. Scientific Reports
\textbf{6}(1): 27234. Nature Publishing Group.
doi:\href{https://doi.org/10.1038/srep27234}{10.1038/srep27234}.

\bibitem[\citeproctext]{ref-Funes2020}
Funes, M. 2020, March. Efectos de la pesca de arrastre sobre la
estructura tr{ó}fica del norte del {Golfo San Jorge}. PhD thesis.

\bibitem[\citeproctext]{ref-Funes2019}
Funes, M., Marinao, C., and Galván, D.E. 2019. Does trawl fisheries
affect the diet of fishes? {A} stable isotope analysis approach*.
Isotopes in Environmental and Health Studies \textbf{55}(4): 327--343.
Taylor \& Francis.
doi:\href{https://doi.org/10.1080/10256016.2019.1626381}{10.1080/10256016.2019.1626381}.

\bibitem[\citeproctext]{ref-Funes2022}
Funes, M., Saravia, L.A., Cordone, G., Iribarne, O.O., and Galván, D.E.
2022. Network analysis suggests changes in food web stability produced
by bottom trawl fishery in {Patagonia}. Scientific Reports
\textbf{12}(1): 10876. Nature Publishing Group.
doi:\href{https://doi.org/10.1038/s41598-022-14363-y}{10.1038/s41598-022-14363-y}.

\bibitem[\citeproctext]{ref-Gaitan2016}
Gaitán, E., and Marí, N. 2016. {An{á}lisis de las comunidades
bent{ó}nicas asociadas a capturas de la flota comercial dirigida a
Macruronus magellanicus}. Instituto Nacional de Investigaci{ó}n y
Desarrollo Pesquero (INIDEP).

\bibitem[\citeproctext]{ref-Galvan2022}
Galván, D.E., Bovcon, N.D., Cochia, P.D., González, R.A., Lattuca, M.E.,
Reinaldo, M.O., Rincón-Díaz, M.P., Romero, M.A., Vanella, F.A., Venerus,
L.A., and Svendsen, G.M. 2022. Changes in the {Specific} and
{Biogeographic Composition} of {Coastal Fish Assemblages} in
{Patagonia}, {Driven} by {Climate Change}, {Fishing}, and {Invasion} by
{Alien Species}. \emph{In} Global {Change} in {Atlantic Coastal
Patagonian Ecosystems}: {A Journey Through Time}. \emph{Edited by} E.W.
Helbling, M.A. Narvarte, R.A. González, and V.E. Villafañe. Springer
International Publishing, Cham. pp. 205--231.
doi:\href{https://doi.org/10.1007/978-3-030-86676-1_9}{10.1007/978-3-030-86676-1\_9}.

@@

\bibitem[\citeproctext]{ref-Garcia-Borboroglu2008}
García-Borboroglu, P., Boersma, P.D., Reyes, L., and Skewgar, E. 2008.
Petroleum {Pollution} and {Penguins}: {Marine Conservation Tools} to
{Reduce} the {Problem}. \emph{In} Marine {Pollution}: {New} research,
Hofer, T.N. Nova Science Publishers Inc., New York, USA. pp. 339--356.

@@

\bibitem[\citeproctext]{ref-Gellner2023}
Gellner, G., McCann, K., and Hastings, A. 2023. Stable diverse food webs
become more common when interactions are more biologically constrained.
Proceedings of the National Academy of Sciences \textbf{120}(31):
e2212061120. Proceedings of the National Academy of Sciences.
doi:\href{https://doi.org/10.1073/pnas.2212061120}{10.1073/pnas.2212061120}.

@@

\bibitem[\citeproctext]{ref-Gongora2012}
Góngora, M.E., González-Zevallos, D., Pettovello, A., and Mendía, L.
2012. Caracterizaci{ó}n de las principales pesquer{í}as del golfo {San
Jorge Patagonia}, {Argentina}. Latin american journal of aquatic
research \textbf{40}(1): 1--11. Pontificia Universidad Cat{ó}lica de
Valpara{í}so. Facultad de Recursos Naturales. Escuela de Ciencias del
Mar. {[}accessed 14 August 2023{]}.

@@

\bibitem[\citeproctext]{ref-Gorini2021}
Gorini, F.L., Lukaszewicz, G., and Giussi, A.R. 2021. {Actualizaci{ó}n
de la estad{í}stica pesquera de peces demersales australes en el
Atl{á}ntico Sudoccidental (per{í}odo 2008-2020)}. \{Informe
T\{\textbackslash\textquotesingle e\}cnico Oficial\}, Instituto Nacional
de Investigaci{ó}n y Desarrollo Pesquero (INIDEP), Mar del Plata.

\bibitem[\citeproctext]{ref-Guihou2020}
Guihou, K., Piola, A.R., Palma, E.D., and Chidichimo, M.P. 2020.
Dynamical connections between large marine ecosystems of austral {South
America} based on numerical simulations. Ocean Science \textbf{16}(2):
271--290. Copernicus GmbH.
doi:\href{https://doi.org/10.5194/os-16-271-2020}{10.5194/os-16-271-2020}.

@@

\bibitem[\citeproctext]{ref-Kawaguchi2009}
Kawaguchi, S., Nicol, S., and Press, A. 2009. Direct effects of climate
change on the {Antarctic} krill fishery. University Of Tasmania.
{[}accessed 13 September 2023{]}.
